---
title: "Downstream analysis followig the image analysis to quanttify Dvl2_mEos codensates (Figure S8C)"
author:  
- name: "Christian Scheeder"
- name: "Antionia Schubert"
  affiliation:
  - "Division Signalling and Functional Genomics, German Cancer Research Center (DKFZ) Heidelberg"  
date: \today
output:
    BiocStyle::pdf_document
vignette: >
    %\VignetteIndexEntry{Vignette Title}
    %\VignetteEngine{knitr::rmarkdown}
editor_options: 
    chunk_output_type: console
---


# Load dependencies
```{r setup, message=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(here)
library(cowplot)
library(ggpubr)
theme_set(theme_cowplot())
```


# Preamble
The image analysis was carried out as follows:
In a first step cells were identified by segmentation. Nuclei were identified 
based on a Hoechst staining (DAPI channel) and cell bodies 
based on a DyLight Phalloidin staining 
(Cy5 channel). 
After segmentation condensates inside cells were identified based on intensity-based 
thresholding (FITC channel) and counted. Condensate counts and cell counts (number of segmented cells) were saved for each image.
An exemplary image analysis can be found with the script
images_analysis_condesate_quantification.Rmd in this repository.
Two different cell lines and different conditions were compared.  
Cell lines:
- HEK cells with a DVL1-mEOS tag
- HEK cells with a DVL1-mEOS tag and a EVI k.o. (clone 1)

Conditions:
- transfection reagent control (RNAiMax)
- transfection reagent (RNAiMax) + non-targeting siRNA
- transfection reagent (RNAiMax) + siCTNNB1


  
   
The time course experiment 4 separate xcperiments (i.e. on different days) 
were performed. Each experiment was carried out on one 384-well plates. Per 
condition, 42 wells with 4 fields of view per well were analyzed. 

# Load the data after the image analysis
Note: The raw data for Figure S8C was saved in one data frame with 
the war data for figure 4G.
```{r}
raw_data <- read_rds(here("raw_data/raw_data_Fig4G","raw_data_Fig4G.rds"))
```


# Prepare boxplot
```{r}


boxplot <- raw_data %>% filter(plate_id == "Plate_IV") %>% 
  filter(treatment %in% c("RNAiMax","siCTNNB1_PoolDharm","siCtl")) %>% 
   mutate(norm_punct_count = (punct_count_adapt/cell_count)*100) %>%
    group_by(barcode,cell_line,plate_id,experiment,treatment) %>%
    summarise_if(is.numeric,mean) %>% 
    ungroup() %>% 
       unite(condition,cell_line,treatment,sep=":") %>% 
        arrange(condition) %>% 
        ggplot(aes(x = condition,y=norm_punct_count)) + geom_boxplot(outlier.shape = NA) + 
       geom_point(aes(color = experiment),
                  position=position_jitterdodge(
                    jitter.width=0, dodge.width = 0.3, seed = 1234), 
             size = 2) +
    ylab("condensate count per 100 cells") +
    xlab("condition") + 
    ylim(0,3.5) +
    theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))

   

boxplot


ggsave(here("./graphics/FigureS8C/","boxplot_figures8c_raw.pdf"),
            plot=boxplot,
            width = 6,height = 8)


```


# Session info
```{r}
sessionInfo()
```