---
title: "Downstream analysis followig the image analysis to quanttify Dvl2_mEos codensates"
author:  
- name: "Christian Scheeder"
- name: "Antionia Schubert"
  affiliation:
  - "Division Signalling and Functional Genomics, German Cancer Research Center (DKFZ) Heidelberg"  
date: \today
output:
    BiocStyle::pdf_document
vignette: >
    %\VignetteIndexEntry{Vignette Title}
    %\VignetteEngine{knitr::rmarkdown}
editor_options: 
    chunk_output_type: console
---


# Load dependencies
```{r setup}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(here)
library(cowplot)
library(ggpubr)
theme_set(theme_cowplot())
```


# Preamble
The image analysis was carried out as follows:
In a first step cells were identified by segmentation. Nuclei were identified 
based on a Hoechst staining (DAPI channel) and cell bodies 
based on a DyLight Phalloidin staining 
(Cy5 channel). 
After segmentation condensates inside cells were identified based on intensity-based 
thresholding (FITC channel) and counted. Condensate counts and cell counts (number of segmented cells) were saved for each image.
An exemplary image analysis can be found with the script
XXX.R in this repository.
Four different cell lines were compared: 
- HEK cells with a DVL1-mEOS tag
- HEK cells with a DVL1-mEOS tag and a EVI k.o. (clone 1)
- HEK cells with a DVL1-mEOS tag and a EVI k.o. (clone 3)
- HEK cells with a DVL1-mEOS tag and a FZD1 k.o. (clone 1)
- HEK wild-type 

For each cell line three replicates were carried out (E01,E02,E03). Per 
replicate multiple 384-well plates were seeded, fixed/stained and imaged. 
For each replicate representative images were selected. The 
image-level data (see description above) was further averaged per plate and
differences between the cell lines were compared using a box plot and statistical 
testing with the ggbbubr package (non-parametric Wilcoxon rank-sum test). 


# Load the data after the image analysis

For each replicate the results were saved in a separate data frame. 

```{r}
results_path <- here("./raw_data/results_image_analysis/data_frames")
files <- dir(results_path, pattern = "*.rds")

raw_data <- files %>% 
  map(~ read_rds(file.path(results_path, .))) %>% 
  reduce(rbind) %>% 
    mutate(cell_line=factor(cell_line,
                            levels = c("wt","mEos","Evi1","Evi3","mFZD")))
```


# Prepare boxplot
```{r}
boxplot <- raw_data %>%   
  separate(well_id,c("well","field")) %>% 
  mutate(norm_punct_count = punct_count_dapt/(cell_count/100)) %>% 
  group_by(experiment,plate_id,cell_line) %>%
  summarise_if(is.numeric,mean) %>%
  ungroup() %>%
    ggplot(aes(x = cell_line,y=norm_punct_count)) + 
    geom_boxplot() + 
    geom_jitter() +
    stat_compare_means(comparisons = list(c("Evi1","mEos"),
                                        c("Evi3","mEos"),
                                        c("mFZD","mEos")),
                       method = "wilcox.test") +
    ylab("normalized condensate count") +
    xlab("cell line")


boxplot

ggsave(here("graphics/Figure4F","boxplot_figure4f_raw.pdf"),
       plot=boxplot,
       width = 6,
       height = 4)
```


# Session info
```{r}
sessionInfo()
```