---
title: "Image analysis to quantify Dvl2_mEos codensates"
author:  
- name: "Christian Scheeder"
- name: "Antionia Schubert"
  affiliation:
  - "Division Signalling and Functional Genomics, German Cancer Research Center (DKFZ) Heidelberg"  
date: \today
output:
    BiocStyle::pdf_document
vignette: >
    %\VignetteIndexEntry{Vignette Title}
    %\VignetteEngine{knitr::rmarkdown}
editor_options: 
    chunk_output_type: console
---


# Load dependencies
```{r setup, message=F}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(fig.align="center")
knitr::opts_chunk$set(dev="png")
library(EBImage)
library(tidyverse)
library(here)
```


# Preamble
Script to document the image analysis carried out to quantify condensates.
In brief, The image analysis was carried out as follows:
In a first step cells were identified by segmentation. Nuclei were identified 
based on a Hoechst staining (DAPI channel) and cell bodies 
based on a DyLight Phalloidin staining 
(Cy5 channel). 
After segmentation condensates inside cells were identified based on intensity-based 
thresholding (FITC channel) and counted. Condensate counts and cell counts (number of segmented cells) were saved for each image.
Four different cell lines were compared: 
- HEK cells with a DVL1-mEOS tag
- HEK cells with a DVL1-mEOS tag and a EVI k.o. (clone 1)
- HEK cells with a DVL1-mEOS tag and a EVI k.o. (clone 3)
- HEK cells with a DVL1-mEOS tag and a FZD1 k.o. (clone 1)
- HEK wild-type  
The image analysis shown here is exemplary carried out for two cell lines (DVL1-mEOS and DVL1-mEOS_EVIk.o._clone1).




# Load the exmaple images 
```{r fig.align="center"}
source_dir_DVL1_mEOS <- 
  here("./raw_data/example_images/HC10984t_HC1098_E01_Evi1_1_2020.10.21.23.32.00")

source_dir_DVL1_Evi1 <- 
  here("./raw_data/example_images/HC10984t_HC1098_E01_mEos_2_2020.10.21.15.56.44")

images_DVL1_mEOS <- EBImage::combine(
  readImage(
    file.path(source_dir_DVL1_mEOS,
              list.files(
                source_dir_DVL1_mEOS)[grepl("DAPI",
                                            list.files(source_dir_DVL1_mEOS))])),
  readImage(
    file.path(source_dir_DVL1_mEOS,
              list.files(
                source_dir_DVL1_mEOS)[grepl("Cy5",
                                            list.files(source_dir_DVL1_mEOS))])),
  readImage(
    file.path(source_dir_DVL1_mEOS,
      list.files(
        source_dir_DVL1_mEOS)[grepl("FITC",
                                    list.files(source_dir_DVL1_mEOS))]))
  )


display(normalize(images_DVL1_mEOS),
        method = "raster",
        all=T)

images_DVL1_Evi1 <- EBImage::combine(
  readImage(
    file.path(source_dir_DVL1_Evi1,
              list.files(
                source_dir_DVL1_Evi1)[grepl("DAPI",
                                            list.files(source_dir_DVL1_Evi1))])),
  readImage(
    file.path(source_dir_DVL1_Evi1,
              list.files(
                source_dir_DVL1_Evi1)[grepl("Cy5",
                                            list.files(source_dir_DVL1_Evi1))])),
  readImage(
    file.path(source_dir_DVL1_Evi1,
      list.files(
        source_dir_DVL1_Evi1)[grepl("FITC",
                                    list.files(source_dir_DVL1_Evi1))]))
  )


display(normalize(images_DVL1_Evi1),
        method = "raster",
        all=T)
```


# Perform image analysis 
```{r fig.align="center"}

# blur the images
images_DVL1_mEOS_smooth <- images_DVL1_mEOS

images_DVL1_mEOS_smooth[,,1] <- gblur(images_DVL1_mEOS_smooth[,,1],
                                      radius = 51, 
                                      sigma = 1)

images_DVL1_mEOS_smooth[,,2] <- gblur(images_DVL1_mEOS_smooth[,,2],
                                      radius  = 51,
                                      sigma = 4)

images_DVL1_mEOS_smooth[,,3] <- gblur(images_DVL1_mEOS_smooth[,,3],
                                      radius  = 51,
                                      sigma = 4)


images_DVL1_Evi1_smooth <- images_DVL1_Evi1

images_DVL1_Evi1_smooth[,,1] <- gblur(images_DVL1_Evi1[,,1],
                                      radius = 51,
                                      sigma = 1)

images_DVL1_Evi1_smooth[,,2] <- gblur(images_DVL1_Evi1[,,2],
                                      radius  = 51,
                                      sigma = 4)

images_DVL1_Evi1_smooth[,,3] <- gblur(images_DVL1_Evi1[,,3],
                                      radius  = 51,
                                      sigma = 4)



# segment the nuclei 
segmentNuclei <- function(inputImage){
  nucleusTresh = thresh(inputImage, 
                        w = 20, h = 20, 
                        offset = 0.004)

  nucleusTresh = fillHull(opening(nucleusTresh,
                                  kern=makeBrush(9, shape="disc")))

  nucleusFill = fillHull(thresh(inputImage, 
                                w = 30, h = 30, 
                                offset = 0.001))

  nucleusRegions = propagate(inputImage,
                             seed=bwlabel(nucleusTresh),
                             mask=nucleusFill)
  
  return(nucleusRegions)
  
}


nuclei_DVL1_mEOS <- segmentNuclei(inputImage=images_DVL1_mEOS_smooth[,,1])
nuclei_DVL1_Evi1 <- segmentNuclei(inputImage=images_DVL1_Evi1_smooth[,,1])

display(
  EBImage::combine(
    colorLabels(nuclei_DVL1_mEOS),
    colorLabels(nuclei_DVL1_Evi1)
  ),
  all=T,
  method = "raster"
  )


# Get outlines of cell bodies
getCellBodies <- function(inputImage,nucleiMask){
   cytoplasmThresh = thresh(inputImage,
                            w = 28,
                            h = 28,
                            offset =  0.001)
      
    cytoplasmOpening = opening(cytoplasmThresh,
                               kern=makeBrush(9,shape="disc"))
      
    cytoplasmOpening2 = opening(inputImage > 0.025)
      
      
    nucleusRegions2 <- nucleiMask
    nucleusRegions2[nucleiMask != 0] <- 1
      
      
    cytoplasmCombined = cytoplasmOpening | cytoplasmOpening2  | nucleusRegions2  
      
      
    storage.mode(cytoplasmCombined) = "integer"
      
      
    cytoplasmRegions = propagate(x = inputImage,
                                 seeds = nucleiMask,
                                 lambda=1e-04,
                                 mask=cytoplasmCombined)
      
    cytoplasmRegions = fillHull(cytoplasmRegions)
    
    return(cytoplasmRegions)
}

cellBodies_DVL1_mEOS <- getCellBodies(inputImage=images_DVL1_mEOS_smooth[,,2],
                                      nucleiMask = nuclei_DVL1_mEOS)

cellBodies_DVL1_Evi1 <- getCellBodies(inputImage=images_DVL1_Evi1_smooth[,,2],
                                      nucleiMask = nuclei_DVL1_Evi1)


display(
  EBImage::combine(
    colorLabels(cellBodies_DVL1_mEOS),
    colorLabels(cellBodies_DVL1_mEOS)
  ),
  all=T,
  method = "raster"
  )


# display the fluorescence image as normalized and pseudo-colored RGB
#   with segmentation borders painted 
ImgColor_DVL1_mEOS = rgbImage(2*normalize(images_DVL1_mEOS_smooth[,,2]),
                    2*normalize(images_DVL1_mEOS_smooth[,,3]),
                    2*normalize(images_DVL1_mEOS_smooth[,,1]))

ImgOut_DVL1_mEOS = paintObjects(cellBodies_DVL1_mEOS,
                      paintObjects(nuclei_DVL1_mEOS,
                                   ImgColor_DVL1_mEOS,
                                   col='yellow'),
                      col='white')


ImgColor_DVL1_Evi1 = rgbImage(2*normalize(images_DVL1_Evi1_smooth[,,2]),
                    2*normalize(images_DVL1_Evi1_smooth[,,3]),
                    2*normalize(images_DVL1_Evi1_smooth[,,1]))

ImgOut_DVL_Evi1 = paintObjects(cellBodies_DVL1_Evi1,
                      paintObjects(nuclei_DVL1_Evi1,
                                   ImgColor_DVL1_Evi1,
                                   col='yellow'),
                      col='white')

display(
  EBImage::combine(
    ImgOut_DVL1_mEOS,
    ImgOut_DVL_Evi1
  ),
  all=T,
  method = "raster"
  )


# Detect mEos condensates in images by adaptive thresholding
getCondensates <- function(inputImage,cellBodyMask){
  condensates_adaptTresh <- thresh(inputImage, 
                                   w = 20, h = 20, 
                                   offset = 0.006)
  
  condensates_adaptTresh[cellBodyMask == 0] <- 0
  
  condensates_adaptTresh_out <- colorLabels(bwlabel(condensates_adaptTresh))
  
  return(condensates_adaptTresh_out)
}


DVL1_mEOS_conTresh <- getCondensates(inputImage = images_DVL1_mEOS_smooth[,,3],
                                     cellBodyMask = cellBodies_DVL1_mEOS)

DVL1_Evi1_conTresh <- getCondensates(inputImage = images_DVL1_Evi1_smooth[,,3],
                                     cellBodyMask = cellBodies_DVL1_Evi1)


display(
  EBImage::combine(
    DVL1_mEOS_conTresh,
    DVL1_Evi1_conTresh    
  ),
  method = 'raster',
  all=T
)


# the number of condensates corresponds to the number of objects in 
# the binary image with labels 

print(paste0(
  "Number of detected condensates in the image with DVL1-mEOS cells: ",
  length(unique(bwlabel(DVL1_mEOS_conTresh)[bwlabel(DVL1_mEOS_conTresh)!=0])),
  " in ",
  length(unique(cellBodies_DVL1_mEOS[cellBodies_DVL1_mEOS!=0])),
  " cells."
  )
)

print(paste0(
  "Number of detected condensates in the image with DVL1-mEOS Evi1 k.o. cells: ",
  length(unique(bwlabel(DVL1_Evi1_conTresh)[bwlabel(DVL1_Evi1_conTresh)!=0])),
  " in ",
  length(unique(cellBodies_DVL1_Evi1[cellBodies_DVL1_Evi1!=0])),
  " cells."
  )
)

```

# Session info
```{r}
sessionInfo()
```